---
title: 'Sentiment Analysis Pride and Prejudice'
date: "created on 22 November 2020 and updated 25-05-2026"
author: "by Misja Bjørk Lauridsen"
output: html_document
---
#quick thoughts
I am working to do a sentiment analysis on Pride and Prejudice by Jane Austen. A well known critic of gentry and society during the regency era of Great Britain. She often places her works within romance settings, as that is the expectation of women in that era. I expect the sentiment to be evenly split. Positive in its romance aspects and negative when describing society.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)

# For text mining:

library(pdftools)
library(tidytext)
library(textdata) 
library(ggwordcloud)
```

### Pride and Prejudice book:
```{r get-document}
paj_path <- here::here("data", "prideandprejudice.pdf")
paj_text <- pdf_text(paj_path)
```

Some things to notice:

#Do you think it will work with any PDF?
#No. Low quality scans of physical books, or pdfs with corruption present in the file will make it difficult or impossible to read the data within the file

```{r single-page}
paj_p9 <- paj_text[9]
paj_p9
```

#See how that compares to the text in the PDF on Page 9. What has pdftools added and where?
#pdftools has added '\n\' for line breaks, and '\' for when an end quotation sign. '\n' when there simply is a new line, but not a line break, such as when a sentence continues onto the next line.

### Some wrangling:

- Split up pages into separate lines (separated by `\n`) using `stringr::str_split()`
- Unnest into regular columns using `tidyr::unnest()`
- Remove leading/trailing white space with `stringr::str_trim()`

```{r split-lines}
paj_df <- data.frame(paj_text)%>%
 mutate(text_full=str_split(paj_text,pattern='\n'))%>%
 unnest(text_full)%>%
 mutate(text_full=str_trim(text_full))
```

Now each line, on each page, is its own row, with extra starting & trailing spaces removed. 

### Get the tokens (individual words) in tidy format

Use `tidytext::unnest_tokens()` (which pulls from the `tokenizer`) package, to split columns into tokens. We are interested in *words*, so that's the token we'll use:

```{r tokenize}
paj_tokens <- paj_df %>% 
     unnest_tokens(word, text_full)
# Each word has its own row!
```

Let's count the words!
```{r count-words}
paj_wc <- paj_tokens %>% 
     count(word) %>% 
     arrange(-n)
paj_wc
```

OK...so we notice that a whole bunch of things show up frequently that we might not be interested in ("a", "the", "and", etc.). These are called *stop words*. Let's remove them. 

### Remove stop words:

See `?stop_words` and `View(stop_words)`to look at documentation for stop words lexicons.

We will *remove* stop words using `tidyr::anti_join()`:
```{r stopwords}
paj_stop <- paj_tokens %>% 
  anti_join(stop_words) %>% 
  select(-paj_text)
```

Now check the counts again: 
```{r count-words2}
paj_swc <- paj_stop %>% 
     count(word) %>% 
     arrange(-n)
```

If i wanted to get rid of the numbers in my text, I could do this:
```{r skip-numbers}
# This code will filter out numbers by asking:
# If you convert to as.numeric, is it NA (meaning those words)?
# If it IS NA (is.na), then keep it (so all words are kept)
# Anything that is converted to a number is removed

paj_no_numeric <- paj_stop %>% 
  filter(is.na(as.numeric(word)))
```
### A word cloud of Pride and Prejudice words (non-numeric)

```{r wordcloud-prep}
length(unique(paj_no_numeric$word))
#This tells us that pride and prejudice has 6114 unique words i.e. words no on the stopwords list and not numerical in category
# We probably don't want to include them all in a word cloud. Let's filter to only include the top 100 most frequent?
paj_top100 <- paj_no_numeric %>% 
  count(word) %>% 
  arrange(-n) %>% 
  head(100)
paj_top100
```
Now I am making the word cloud
```{r wordcloud}
paj_cloud <- ggplot(data = paj_top100, aes(label = word)) +
  geom_text_wordcloud() +
  theme_minimal()

paj_cloud
```

That's underwhelming. Let's customize it a bit: I am choosing the colours be the popular colours of fabrics during the regency for the gentry and nobility
```{r wordcloud-pro}
ggplot(data = paj_top100, aes(label = word, size = n)) +
  geom_text_wordcloud_area(
    aes(color = n),
    mask = png::readPNG(
      system.file("extdata/hearth.png",
                  package = "ggwordcloud",
                  mustWork = TRUE)
    ),
    rm_outside = TRUE
  ) +
  scale_size_area(max_size = 12) +
  scale_color_gradientn(colors = c("pink", "purple", "skyblue")) +
  theme_minimal()
```
Cute! However the heartshape is lost upon being exported ':('

### Sentiment analysis

The three general-purpose lexicons are

  -  AFINN from Finn Årup Nielsen,
  -  bing from Bing Liu and collaborators, and
  -  nrc from Saif Mohammad and Peter Turney

Let's explore the sentiment lexicons.

"afinn": Words ranked from -5 (very negative) to +5 (very positive)
```{r afinn}
get_sentiments(lexicon = "afinn")


# Let's look at the pretty positive words:
afinn_pos <- get_sentiments("afinn") %>% 
  filter(value %in% c(3,4,5))

afinn_pos
```

bing: binary, "positive" or "negative"
```{r bing}
get_sentiments(lexicon = "bing")

```

nrc:https://saifmohammad.com/WebPages/NRC-Emotion-Lexicon.htm
Includes bins for 8 emotions (anger, anticipation, disgust, fear, joy, sadness, surprise, trust) and positive / negative. 

**Citation for NRC lexicon**: Crowdsourcing a Word-Emotion Association Lexicon, Saif Mohammad and Peter Turney, Computational Intelligence, 29 (3), 436-465, 2013.

Now nrc:
```{r nrc}
get_sentiments(lexicon = "nrc")

```

Let's do sentiment analysis on the Pride and Prejudice text data using afinn, and nrc. 


### Sentiment analysis with afinn: 

First, bind words in `paj_stop` to `afinn` lexicon:
```{r bind-afinn}
paj_afinn <- paj_stop %>% 
  inner_join(get_sentiments("afinn"))
paj_afinn
```

Let's find some counts (by sentiment ranking):
```{r count-afinn}
paj_afinn_hist <- paj_afinn %>% 
  count(value)

# Plot them: 
ggplot(data = paj_afinn_hist, aes(x = value, y = n)) +
  geom_col(aes(fill = value)) +
  theme_bw()
```

Investigate some of the words in a bit more depth:
```{r afinn-2}
# What are these '2' words?
paj_afinn2 <- paj_afinn %>% 
  filter(value == 2)
```

```{r afinn-2-more}
# Check the unique 2-score words:
unique(paj_afinn2$word)

# Count & plot them
paj_afinn2_n <- paj_afinn2 %>% 
  count(word, sort = TRUE) %>% 
  filter(n > 10) %>% 
  mutate(word = fct_reorder(factor(word), n))

ggplot(data = paj_afinn2_n, aes(x = word, y = n)) +
  geom_col() +
  coord_flip() +
  theme_bw()
#I am filtering unique words that have a usage below 12 incident, simply due to the massive amounts of unique words utilised throughout the entire book.

```

However, proud in this lexicon is utilised positively, when in the book it is used negatively. It is pride *and* prejudice, they're, throughout the boook, compared and put as equals. You're prejudiced because you're proud and you're proud because you're prejudiced. Elizabeth and Darcy must both overcome their Pride, and Elizabeth must certainly overcome her prejudice, for them to get a good ending.

We learn something important from this example: Just using a sentiment lexicon to match words will not differentiate between different uses of the word...(ML can start figuring it out with context, but we won't do that here).

Or we can summarize sentiment for the report: 
```{r summarize-afinn}
paj_summary <- paj_afinn %>% 
  summarize(
    mean_score = mean(value),
    median_score = median(value)
  )
paj_summary 
```

The mean and median indicate *slightly* positive overall sentiments based on the AFINN lexicon. That is a mean score of 0.343 and a median score of 1. Which makes sense it is a romance book.

### NRC lexicon for sentiment analysis

We can use the NRC lexicon to start "binning" text by the feelings they're typically associated with. As above, we'll use inner_join() to combine the Pride and Prejudice non-stopword text with the nrc lexicon: 

```{r bind-bing}
paj_nrc <- paj_stop %>% 
  inner_join(get_sentiments("nrc"))
```

Wait, won't that exclude some of the words in our text? YES! We should check which are excluded using `anti_join()`:

```{r check-exclusions}
paj_exclude <- paj_stop %>% 
  anti_join(get_sentiments("nrc"))

# View(paj_exclude)

# Count to find the most excluded:
paj_exclude_n <- paj_exclude %>% 
  count(word, sort = TRUE)

head(paj_exclude_n)
#this returns that names of the characters within the book
```

Now find some counts: 
```{r count-bing}
paj_nrc_n <- paj_nrc %>% 
  count(sentiment, sort = TRUE)

# And plot them:

ggplot(data = paj_nrc_n, aes(x = sentiment, y = n)) +
  geom_col(aes(fill = sentiment))+
  theme_bw()
```

Or count by sentiment *and* word, then facet:
```{r count-nrc}
paj_nrc_n5 <- paj_nrc %>% 
  count(word,sentiment, sort = TRUE) %>% 
  group_by(sentiment) %>% 
  top_n(5) %>% 
  ungroup()

paj_nrc_gg <- ggplot(data = paj_nrc_n5, aes(x = reorder(word,n), y = n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, ncol = 2, scales = "free") +
  coord_flip() +
  theme_minimal() +
  labs(x = "Word", y = "count")

# Show it
paj_nrc_gg

# Save it
ggsave(plot = paj_nrc_gg, 
       here("figures","paj_nrc_sentiment.png"), 
       height = 8, 
       width = 5)

```

Confidence still registers as a positive trait, even, as we read the book it is painted as a negative trait.
```{r nrc-proud}
conf <- get_sentiments(lexicon = "nrc") %>% 
  filter(word == "proud")

# Yep, check it out:
conf
```


## Takeaway points

It is a quick tool to gauge the overall sentiments and opinions within the text itself, however it will miss finer points being made, this could be troublesome if the user does not know if a book or text is a satire or the like. This could cause wrong impressions of the litterature.
I was right in my assumption that the sentiments within the book would be fairly evenly split.

### Credits: 
This tutorial is inspired by Allison Horst's Advanced Statistics and Data Analysis.
